Nearest Neighbour Methods for Length of Stay Prediction

### 2015-02-03 papers/paper.tex
Finish introduction

In this paper we present the application of nearest neighbour methods
for predicting the length of stay category for patients admitted to a trauma
ward. The categories are \textit{less than or equal to 2 days} or
\textit{greater than 2 days}, which have been used in previous work on length
of stay prediction in the same medical domain \citep{Dinh2013a}.
Nearest neighour methods have not been studied in predicting the length of stay,
and we believe they are much more intuitive to
understand than techniques such as support vector machines.
Additionally, they can be adjusted to suit particular medical domains by
changing their similarity function. We compare the performance of the basic
nearest neighbour algorithm, as well as an extension (K*), against the
performance
of previous work which used logistic regression. We also propose a novel
extension to the standard nearest neighbour algorithm, Ranked Distance, and
test its performance against the other classifiers. In addition, we apply a
range of feature selection methods, as they have not been
thoroughly studied in existing work. Finally, we compare the performance of
all of these with the state-of-the-art support vector machines and the
commonly-used artificial neural networks.

We found that the standard 1-nearest neighbour classifier achieved the highest
accuracy of 77.81\%, an improvement of 2.75\% over previous work using the same
data set carried out by Dinh et al.\citep{Dinh2013a}.

\section{Nearest Neighbour Classification}
In this section we will describe the $k$-nearest neighbour ($k$-NN) algorithm
and discuss some extensions which we use in our work. Additionally, we will
describe our contribution, which is an extension of the distance metric for
finding the nearest neighbours.

@@

\section{Evaluation}
\subsection{Data Set}
Our evaluation was carried out using a trauma-specific data set
containing 2546 records collected from 2007--2011 collected at a major hospital
in Sydney, Australia.

@@

\section{Results and Discussion}
The best accuracy was 77.81\%, achieved using the $k$-NN algorithm
with 1 nearest neighbour, a discretised data set, and feature selection using a
C4.5 wrapper. The best AUC of 0.846 was achieved using logistic regression with
features selected by a correlation coefficient threshold of 0.1, and again
using the discretised data set.

@@

\section{Conclusions}
The best accuracy we achieved on the trauma data was 77.81\% with the $k$-NN
classifier using 1 nearest neighbour and 11 out of 78 features. This represents
an improvement over previous work done by Dinh et al. \cite{Dinh2013a}
on the same data set,
where they managed to achieve 75.06\% with logistic regression. More
importantly, we were able to improve upon their AUC of 0.812 with a result of
0.846, also using logistic regression but first discretising numeric features
and reducing the feature set to 29 features.
If we consider an even smaller number of features, then K*
gives an AUC of 0.838 with 11 features selected by the C4.5 wrapper, which
improves upon the result of Dinh et al. \cite{Dinh2013a} while using a little
more than half
the number of features. The drawback of $k$-NN and K* is that they require all
examples to be stored and then searched during classification, which becomes
storage- and computation-expensive as the data set increases in size.

>>>>

In this paper we present the application of nearest neighbour methods
in predicting the length of stay category for patients admitted to a trauma
ward. The categories are \textit{less than or equal to 2 days} or
\textit{greater than 2 days}, which have been used in previous work on length
of stay prediction in the same medical domain \citep{Dinh2013a}.
Nearest neighour methods have not been studied in predicting the length of stay,
and we believe they are much more intuitive to
understand than techniques such as support vector machines.
Additionally, they can be adjusted to suit particular medical domains by
changing their similarity function. We compare the performance of the basic
nearest neighbour algorithm, as well as an extension (K*), against the
performance
of previous work which used logistic regression. We also propose a novel
extension to the standard nearest neighbour algorithm, Ranked Distance, and
test its performance against the other classifiers. In addition, we apply a
range of feature selection methods and systematically investigate the effect
of discretisation, as they have not been
thoroughly studied in existing work. Finally, we compare the performance of
these with the state-of-the-art support vector machines and the
commonly-used artificial neural networks.

We found that the standard nearest neighbour classifier using only 1 nearest
neighbour achieved a ten-fold cross-validated accuracy of 77.81\%, the highest
out of all the methods. This is an improvement of 2.75\% over previous work
carried out by Dinh et al. \citep{Dinh2013a} using the same data set. Also, the
nearest neighbour classifiers almost always outperformed the logistic
regression, support vector machines and neural networks when the number of
features was gradually reduced. This means that predictions are able to be made
more quickly, and will also be more intuitive to interpret and understand for
physicians.

In Section \ref{sec:nn}, we outline the basics of nearest neighbour
classification and describe both the existing K* extension and our proposed
extension, Ranked Distance. Section \ref{sec:eval} describes the method used
to evaluate the performance of these classifiers, and Section \ref{sec:results}
proceeds to discuss the results. We present concluding remarks in Section
\ref{sec:conclusions}.

\section{Nearest Neighbour Classification}
\label{sec:nn}
In this section we will describe the $k$-nearest neighbour ($k$-NN) algorithm
and discuss some extensions which we use in our work. Additionally, we will
describe our contribution, which is an extension of the distance metric for
finding the nearest neighbours.

@@

\section{Evaluation}
\label{sec:eval}
\subsection{Data Set}
Our evaluation was carried out using a trauma-specific data set
containing 2546 records collected from 2007--2011 collected at a major hospital
in Sydney, Australia.

@@

\section{Results and Discussion}
\label{sec:results}
The best accuracy was 77.81\%, achieved using the $k$-NN algorithm
with 1 nearest neighbour, a discretised data set, and feature selection using a
C4.5 wrapper. The best AUC of 0.846 was achieved using logistic regression with
features selected by a correlation coefficient threshold of 0.1, and again
using the discretised data set.

@@

\section{Conclusions}
\label{sec:conclusions}
The best accuracy we achieved on the trauma data was 77.81\% with the $k$-NN
classifier using 1 nearest neighbour and 11 out of 78 features. This represents
an improvement over previous work done by Dinh et al. \cite{Dinh2013a}
on the same data set,
where they managed to achieve 75.06\% with logistic regression. More
importantly, we were able to improve upon their AUC of 0.812 with a result of
0.846, also using logistic regression but first discretising numeric features
and reducing the feature set to 29 features.
If we consider an even smaller number of features, then K*
gives an AUC of 0.838 with 11 features selected by the C4.5 wrapper, which
improves upon the result of Dinh et al. \cite{Dinh2013a} while using a little
more than half
the number of features. The drawback of $k$-NN and K* is that they require all
examples to be stored and then searched during classification, which becomes
storage- and computation-expensive as the data set increases in size.
